---
title: "Tooth Growth Analyses"
author: "João Pedro Schmitt"
date: "4 de outubro de 2016"
output: html_document
---

For this assigment was used the dataset ToothGrowth.
The CodeBook is:
1. len - numeric - Tooth length
2. supp - factor - Supplement type (VC or OJ)
3. dose - numeric - Dose in milligrams/day

### 1. Load the ToothGrowth data and perform some basic exploratory data analyses

A graph demonstring the relationship of the variables:
```{r message=FALSE}
data(ToothGrowth)

library(ggplot2)
library(dplyr)

filtered <- ToothGrowth %>% 
    group_by(supp, dose) %>% 
    summarise_each(funs(sum(len))) %>% 
    as.data.frame()

filtered$dose <- as.factor(filtered$dose)

names(filtered) <- c("Supplement","Dose",'Length')

g <- ggplot(data = filtered, 
    aes(x = Dose, y = Length))
g <- g + geom_bar(stat = "identity", 
    aes(fill=Supplement), 
    position=position_dodge())
g <- g + labs(title = "ToothGrowth by supplement")
print(g)
```

### 2. Provide a basic summary of the data.

A little summary of the variables:

```{r}
summary(ToothGrowth)
```

### 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.

To calculate the data we use the follow function:

```{r}
calculateIntervals <- function(g1, g2) {
    
    # Get the mean
    g1m <- mean(g1)
    g2m <- mean(g2)
    
    # Get the standard deviation
    g1s <- sd(g1)
    g2s <- sd(g2)
    
    # Get the number of each collection
    g1n <- length(g1)
    g2n <- length(g2)
    
    # Get the degrees of freedom
    df <- (g1s^2/g1n + g2s^2/g2n)^2 / ((g1s^2/g1n)^2/(g1n-1) + (g2s^2/g2n)^2/(g2n-1))
    
    # Calculate the confidence interval for 5% of both sides using T Student's
    int <- (g1m - g2m) + c(-1,1) * qt(.975, df) * sqrt(g1s^2/g1n + g2s^2/g2n)
    
    return(int)
}

calculatePValue <- function(g1, g2) {
    
    # Order to get from the smaller to greater
    if(mean(g1) > mean(g2)) {
        aux <- g1
        g1 <- g2
        g2 <- aux
    }
        
    # Get the mean
    g1m <- mean(g1)
    g2m <- mean(g2)
    
    # Get the standard deviation
    g1s <- sd(g1)
    g2s <- sd(g2)
    
    # Get the number of each collection
    g1n <- length(g1)
    g2n <- length(g2)
    
    # Get the degrees of freedom
    df <- (g1s^2/g1n + g2s^2/g2n)^2 / ((g1s^2/g1n)^2/(g1n-1) + (g2s^2/g2n)^2/(g2n-1))
    
    # Get T statistic
    t <- (g1m - g2m) / sqrt(g1s^2/g1n + g2s^2/g2n)
    
    # Return the p-value
    return(2 * pt(t, df))
    
}
```

#### Classifing by supplement

The confidence intervals are:
```{r}
g1 <- ToothGrowth$len[ToothGrowth$supp == "OJ"]
g2 <- ToothGrowth$len[ToothGrowth$supp == "VC"]
calculateIntervals(g1, g2)
```

The p-value is:
```{r}
g1 <- ToothGrowth$len[ToothGrowth$supp == "OJ"]
g2 <- ToothGrowth$len[ToothGrowth$supp == "VC"]
calculatePValue(g1, g2)
```

For a confidence interval of 5% the p-value was 6%, so the difference by supplement is not significant to consider one supplement instead of the other.

#### Classifing by dose of 0.5 and 1.0

The confidence intervals are:
```{r}
g1 <- ToothGrowth$len[ToothGrowth$dose == 0.5]
g2 <- ToothGrowth$len[ToothGrowth$dose == 1.0]
calculateIntervals(g1, g2)
```

The p-value is:
```{r}
g1 <- ToothGrowth$len[ToothGrowth$dose == 0.5]
g2 <- ToothGrowth$len[ToothGrowth$dose == 1.0]
calculatePValue(g1, g2)
```


#### Classifing by dose of 0.5 and 2.0

The confidence intervals are:
```{r}
g1 <- ToothGrowth$len[ToothGrowth$dose == 0.5]
g2 <- ToothGrowth$len[ToothGrowth$dose == 2.0]
calculateIntervals(g1, g2)
```

The p-value is:
```{r}
g1 <- ToothGrowth$len[ToothGrowth$dose == 0.5]
g2 <- ToothGrowth$len[ToothGrowth$dose == 2.0]
calculatePValue(g1, g2)
```

#### Classifing by dose of 1.0 and 2.0

The confidence intervals are:
```{r}
g1 <- ToothGrowth$len[ToothGrowth$dose == 1.0]
g2 <- ToothGrowth$len[ToothGrowth$dose == 2.0]
calculateIntervals(g1, g2)
```

The p-value is:
```{r}
g1 <- ToothGrowth$len[ToothGrowth$dose == 1.0]
g2 <- ToothGrowth$len[ToothGrowth$dose == 2.0]
calculatePValue(g1, g2)
```