---
title: "Statistical inference course project"
author: "João Pedro Schmitt"
date: "3 de outubro de 2016"
output: html_document
---

```{r}
# Randomness parameter
set.seed(32016)

# Exponential distribution paramenters
lambda <- 0.2
exp_n <- 40

# Exponential distribution simulation
exp_sim_mean <- numeric()
exp_sim_sd <- numeric()
for(i in 1:10000) {
  simulation <- rexp(exp_n, lambda)
  exp_sim_mean <- c(exp_sim_mean, mean(simulation))
  exp_sim_sd <- c(exp_sim_sd, sd(simulation))
}
```

### Question 1

Show the sample mean and compare it to the theoretical mean of the distribution with the means.

Comparing the theorical mean: 
```{r} 
exp_theor_mean <- 1/lambda
exp_theor_mean
``` 
With the mean of the simulations:
```{r}
exp_sim_mean_est <- mean(exp_sim_mean)
exp_sim_mean_est
```
The mean difference between the simulated and theorical is:
```{r}
abs(exp_theor_mean - exp_sim_mean_est)
```
Below follow the historgram of the distribution:
```{r}
hist(exp_sim_mean, 
     main = "Exponential distribuition with lambda equals to 0.2.", 
     xlab = "Mean", 
     col = "tomato")
abline(v = exp_sim_mean_est, 
       col = "yellow", 
       lwd = 4)
abline(v = exp_theor_mean, 
       col = "blue", 
       lwd = 2)
legend("topright", 
       legend = c("Sample mean", "Theorical mean"), 
       col = c("yellow","blue"), 
       lwd = c(2,2))
```

### Question 2

Show how variable the sample is (via variance) and compare it to the theoretical variance of the distribution.

The standard deviation of simulations is equals to:
```{r}
exp_sim_sd_est <- mean(exp_sim_sd)/sqrt(exp_n)
exp_sim_sd_est
```
The standard deviation of theorical is equals to:
```{r}
exp_theor_sd <- 1/lambda/sqrt(exp_n)
exp_theor_sd
```
The standard deviantion of the difference between the simulated and theorical is:
```{r}
abs(exp_theor_sd - exp_sim_sd_est)
```
  
  
The variance of the distribution simulation is:
```{r}
exp_sim_var_est <- exp_sim_sd_est^2
exp_sim_var_est
```
The variance of the distribution theorical is:
```{r}
exp_theor_var <- exp_theor_sd^2
exp_theor_var
```
The variance of the difference between the simulated and theorical is:
```{r}
abs(exp_theor_var - exp_sim_var_est)
```

### Question 3

Show that the distribution is approximately normal.

The follow code demonstra the normal distribuition:
```{r}
hist(exp_sim_mean, 
     main = "Exponential distribuition with lambda equals to 0.2.", 
     xlab = "Mean", 
     col = "tomato", 
     prob = T)
lines(density(exp_sim_mean), 
      col="green", 
      lwd = 2)
```
