---
title: "Tooth Growth Analyses"
author: "João Pedro Schmitt"
date: "4 de outubro de 2016"
output: html_document
---

For this assigment was used the dataset ToothGrowth.
The CodeBook is:
1. len - numeric - Tooth length
2. supp - factor - Supplement type (VC or OJ)
3. dose - numeric - Dose in milligrams/day

### 1. Load the ToothGrowth data and perform some basic exploratory data analyses

A graph demonstring the relationship of the variables:
```{r message=FALSE}
data(ToothGrowth)

library(ggplot2)
library(dplyr)

filtered <- ToothGrowth %>% 
    group_by(supp, dose) %>% 
    summarise_each(funs(sum(len))) %>% 
    as.data.frame()

filtered$dose <- as.factor(filtered$dose)

names(filtered) <- c("Supplement","Dose",'Length')

g <- ggplot(data = filtered, 
    aes(x = Dose, y = Length))
g <- g + geom_bar(stat = "identity", 
    aes(fill=Supplement), 
    position=position_dodge())
g <- g + labs(title = "ToothGrowth by supplement")
print(g)
```

### 2. Provide a basic summary of the data.

A little summary of the variables:

```{r}
summary(ToothGrowth)
```

### 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.


#### Classifing by supplement

The confidence intervals and the p-value of the hypothesis test by supplement are:

```{r}
t.test(len ~ supp, data = ToothGrowth, paired = F, var.eq = F)
```

For a confidence interval of 5% the p-value was 6%, so the difference by supplement is not significant to consider one supplement instead of the other.

#### Classifing by dose of 0.5 and 1.0

The confidence intervals and the p-value of the hypothesis test by dose of 0.5 and 1.0 are:

```{r}
filtered <- ToothGrowth %>% filter(dose %in% c(0.5,1.0))
t.test(len ~ dose, data = filtered, paired = F, var.eq = F)
```


#### Classifing by dose of 0.5 and 2.0

The confidence intervals and the p-value of the hypothesis test by dose of 0.5 and 1.0 are:

```{r}
filtered <- ToothGrowth %>% filter(dose %in% c(0.5,2.0))
t.test(len ~ dose, data = filtered, paired = F, var.eq = F)
```

#### Classifing by dose of 1.0 and 2.0

The confidence intervals and the p-value of the hypothesis test by dose of 0.5 and 1.0 are:

```{r}
filtered <- ToothGrowth %>% filter(dose %in% c(1.0,2.0))
t.test(len ~ dose, data = filtered, paired = F, var.eq = F)
```